---
title: "NEO-PP-187 (temporary)"
author: "JChoi"
date: "06-11-2020"
output: html_document
---

> This document is for estimating glucose-lowering medicatino effect in diabetes incident cases in the NEO study. The protocol can be found [here: protocol](LINK REMOVED)
> + This is a temporary version to share syntax only.
> + We combine NEO baselie, GP visit, laboratory data.
> + A list for incident diabetes cases is preliminary.


# We take the following steps: 
### Step 1: Prepare datasets (baseline, lab, medication data). Clean up & merge
### Step 2**: Descriptive statistics & plots
### Step 3**: Fitting LMMs
### Step 4**: Fitting spline models
### Step 5**: Check medication information 


```{r echo=FALSE, include=FALSE, message=FALSE}
#setwd("//vf-i-epid/epid$/Choi/NEO-PP-187/syntax")
```

```{r echo=FALSE, warning=FALSE, include=FALSE, message=FALSE}
#install.packages("readstata13")
#install.packages("table1")
#install.packages("ggplot")
#install.packages("lme4")
#install.packages("gt")
#install.packages("ciTools")
#install.packages("rms")
#install.packages("Rfast")
library(readstata13)
library(table1)
library(ggplot2)
library(lme4)
library(dplyr)
library(readr)
library(gt)
library(rms)
library(splines)
library(base)
#library(Rfast)
```

<br>

<br>


# Step 1
### 1.1 Load data files
  + lab: laboratory measurements (hba1c, glucose)
  + med: GP medication data
  + base: NEO baseline data

```{r}
# Data will not be uploaded to github due to privacy
lab<-data.frame(read.dta13("//vf-i-epid/epid$/Choi/NEO-PP-187/LAB_Choi_2020-07-30.dta", nonint.factors=TRUE))
med<-data.frame(read.dta13("//vf-i-epid/epid$/Choi/NEO-PP-187/MEDICATIE_Choi_2020-07-30.dta", nonint.factors=TRUE))
base<-data.frame(read.dta13('//vf-i-epid/epid$/Choi/NEO-PP-187/NEO_analyse_PP187_gluclowmed_Choi_2020-07-30.dta', nonint.factors=TRUE))
#(check what's happening with 'income' var in the baseline data)-> fixed

#View(lab)
#View(med)
#View(base)

head(base)
head(med)
head(lab)
```

<br>

### 1.2 Check unique cases in each dataset
```{r}
#ID's are not coded as number
base$PP187ID<-as.numeric(base$PP187ID )
med$PP187ID<- as.numeric(med$PP187ID)
lab$PP187ID<- as.numeric(lab$PP187ID)

print("# unique IDs in baseline data")
length(unique(base$PP187ID))

print("# unique IDs in medication data")
length(unique(med$PP187ID))

print("# unique IDs in laboratory data")
length(unique(lab$PP187ID))
```

<br>

### 1.3 Identify the IDs included in all three data
```{r}
#base$PP187ID %>% intersect(med$PP187ID) %>% intersect(lab$PP187ID) %>% length
idlist<-base$PP187ID %>% 
          intersect(med$PP187ID) %>% 
          intersect(lab$PP187ID) %>% 
          data.frame

colnames(idlist)[1]<-"PP187ID"
idlist
```

<br>

### 1.4 Select IDs included in 'idlist' in each dataset
```{r}
base.comp<-right_join(base,idlist, by="PP187ID")
lab.comp<-right_join(lab,idlist, by="PP187ID")
med.comp<-right_join(med,idlist, by="PP187ID")

```

<br>

### 1.5 Prepare 'lab' data
  + Cleaning up same measurements coded under different names. 
  + Remove physiologically abnormal values (eg., Hba1c DCCT < 4, glucose = 0) 
```{r}
str(lab.comp)
```

<br>

**1) HbA1C**
  + Recode variables
  + Convert HbA1c measurements (HbA1c was measured in two different ways: DCCT and IFCC).
  + Check abnormal HbA1c values
  + Some ppl have values under 20: set them to missing
  
```{r}
#Change data var. format
lab.comp$testdaten<-lab.comp$testdaten %>% 
                      as.character %>% 
                      as.Date(format="%Y%m%d")


#group HbA1c IFCC measurements together
lab.comp$descriptio[lab.comp$descriptio=="HbA1c (glycohemoglobine) IFCC"] <-"HbA1c IFCC"
lab.comp$descriptio[lab.comp$descriptio=="GlycoHb/HbA1c IFCC"] <-"HbA1c IFCC"
lab.comp$descriptio[lab.comp$descriptio=="HbA1C (nieuw)"] <-"HbA1c IFCC"


#group HbA1c DCCT measurements together
lab.comp$descriptio[lab.comp$descriptio=="glycoHb/HbA1c"] <-"HbA1c DCCT"
lab.comp$descriptio[lab.comp$descriptio=="glycohemoglobine (HbA1c) DCCT"] <-"HbA1c DCCT"


#convert HbA1c DCCT to IFCC
DtoF<-function(x){
  round((x*0.915/10)+2.15,1)
}

lab.comp$valuen2<-ifelse(lab.comp$descriptio=="HbA1c IFCC", 
                         DtoF(lab.comp$valuen), 
                         lab.comp$valuen)

lab.comp$descriptio2<-ifelse(lab.comp$descriptio=="HbA1c IFCC", 
                             "HbA1c DCCT", 
                             lab.comp$descriptio)

#Check abnormal values, set them to missing for now.
#sum(lab.comp$valuen2[lab.comp$descriptio=="HbA1c IFCC"]<20, na.rm=TRUE)
lab.comp$valuen2[lab.comp$valuen2<4 & lab.comp$descriptio2=='HbA1c DCCT']<-NA

```

<br>

  + Plots: HbA1c DCCT before & after cleaning
```{r}
par(mfrow=c(1,2))
hist(lab.comp$valuen[lab.comp$descriptio=="HbA1c DCCT"],
     xlab = "HbA1c DCCT [%]",
     main = "Before cleaning")


hist(lab.comp$valuen2[lab.comp$descriptio=="HbA1c DCCT"],
     xlab = "HbA1c DCCT [%]",
     main = "After cleaning")
```

<br>

**2) Glucose nuchter**
  + Checking physiologically impossible glucose values
```{r}
#range(lab.comp$valuen[lab.comp$descriptio2=='Glucose nuchter'], na.rm=TRUE)

sum(lab.comp$valuen[lab.comp$descriptio2=='Glucose nuchter']>30, na.rm=TRUE)

sum(lab.comp$valuen2[lab.comp$descriptio2=="Glucose nuchter"]==0, na.rm=TRUE)
```
<br>

  + 21 ppl with glucose level >80. (The rest has glucose value under 30): Divide these values by 10, according to Olaf.
  + If glucose level is 0, set it to missing.
```{r}
# Recode names
lab.comp$descriptio2[lab.comp$descriptio=="glucose nuchter, veneus (lab)"] <-"Glucose nuchter"
lab.comp$descriptio2[lab.comp$descriptio=="glucose nuchter, draagbare meter"] <-"Glucose nuchter"

# Unrealistic high values-> divid with 10
lab.comp$valuen2<-ifelse(lab.comp$descriptio2=='Glucose nuchter' & lab.comp$valuen2>30, 
                         lab.comp$valuen2/10,
                         lab.comp$valuen2)

# glucose =0 -> missing
lab.comp$valuen2[lab.comp$valuen2==0 & lab.comp$descriptio2=='Glucose nuchter']<-NA
```

<br>

  + Plots: glucose level before & after cleaning
```{r}
par(mfrow=c(1,2))
hist(lab.comp$valuen[lab.comp$descriptio2=="Glucose nuchter"], 
     xlab = "Fasting glucose [mmol/L]", 
     main="Before cleaning")

hist(lab.comp$valuen2[lab.comp$descriptio2=="Glucose nuchter"], 
     xlab = "Fasting glucose [mmol/L]", 
     main="After cleaning")
```

<br>

### 1.6 Prepare 'base' data
```{r}
#Some factor vars are numeric.
names<- c('sexe',
          'etnizelf',
          'etnwhite',
          'eduh',
          'income',
          'hypertensie',
          'Abdomen',
          'Vetpresent',
          'ATPIIIglucose',
          'glucmet2bl',
          'nuchter',
          'devmeal',
          'medC10liplow',
          "medA10gluclow",
          "medHypertension",
          "medN06A",
          "medN05A",
          "medN05B" ,
          "medH03A",
          "medL02A",
          "medL02B")
base.comp[,names]<-lapply(base.comp[,names],factor)
#str(base)


# Labels the variables, for table 1
label(base.comp$sexe)<-"Sex"
base.comp$sexe<-recode(base.comp$sexe, 'man'='Male', 'vrouw'='Female')
label(base.comp$leeftijd)<-"Age"
label(base.comp$eduh)<-"Educaton"
label(base.comp$hypertensie)<-"Hypertension"
label(base.comp$bmim)<-"BMI"
label(base.comp$glucose1)<-"Glucose"
label(base.comp$Insuline_r1)<-"Insulin"
label(base.comp$homa1IR)<-"HOMA1-IR"
label(base.comp$HBA1C)<-"HbA1c"
label(base.comp$hypertensie)<-"Hypertension"
base.comp$hypertensie<-recode(base.comp$hypertensie, 'nee'='No', 'ja'='Yes')
label(base.comp$choltot1)<-"Total cholesterol"
label(base.comp$trig1)<-"Triglycerides"
label(base.comp$hdlc1)<-"HDL"
label(base.comp$fldl1)<-"LDL"
label(base.comp$medC10liplow)<-"Lipid-lowering drugs use"
label(base.comp$medA10gluclow)<-"Glucose-lowering drugs use"
label(base.comp$medHypertension)<-"Hypertension drugs use"

units(base.comp$leeftijd)<-"years"
units(base.comp$bmim)<-"kg/m2"
units(base.comp$glucose1)<-"mmol/L"
units(base.comp$Insuline_r1)<-"mU/L"
units(base.comp$HBA1C)<-"%"
units(base.comp$choltot1)<-"mmol/L"
units(base.comp$trig1)<-"mmol/L"
units(base.comp$hdlc1)<-"mmol/L"
units(base.comp$fldl1)<-"mmol/L"
```

<br>


### 1.7 Identify the earliest prescription date
  + Identify the earliest prescription date for each individual and defined it as 'time 0'.
```{r}
#Finding the first prescription dates for each person
prescdata_tmp<-matrix(NA, length(base.comp$PP187ID), 2)

for(i in base.comp$PP187ID){
  prescdata_tmp[which(base.comp$PP187ID==i),] <-cbind(i, min(med.comp$prescdate[med.comp$PP187ID==i]))  
  prescdata_tmp
}

prescdata_tmp<-data.frame(prescdata_tmp)
colnames(prescdata_tmp)<-c("PP187ID", "prescdata")
prescdata_tmp
```

<br>


### 1.8 Merge lab & medication data
  + We now know whether each lab measurement is before or after medication prescription & when.
```{r}
#Merge the first prescription dates data to the lab data

lab.med<-merge(lab.comp, prescdata_tmp, by="PP187ID")
lab.med$prescdata <- lab.med$prescdata %>% 
                        as.character %>% 
                        as.Date("%Y%m%d")

#centralize the time to prescription date (in a month-scale)
lab.med$time0<- as.numeric((lab.med$testdaten - lab.med$prescdata)/30.4)

#indicator for before/ after of using medication
lab.med$medindi<-ifelse(lab.med$time0>0, 1, 0)
```

<br>

### 1.9 Creating categorical time variables for the later LMM models
  + Dummy var. for time:
  + 6m to 1y before prescription (reference)
  + 6m to 3m before prescription
  + 3m to 0m before prescription
  + 0m to 3m after prescription
  + 3m to 6m after prescription
  + 6m to 1y after prescription
```{r}
#Categorical time variable
lab.med$timebefore_6m<-ifelse(lab.med$time0 <= -6, 1, 0)
lab.med$time_6m<-ifelse(lab.med$time0 > -6 & lab.med$time0 <= -3, 1, 0)
lab.med$time_3m<-ifelse(lab.med$time0 > -3 & lab.med$time0 <= 0, 1, 0)
lab.med$time3m<-ifelse(lab.med$time0>0 & lab.med$time0<=3, 1, 0)
lab.med$time6m<-ifelse(lab.med$time0>3 & lab.med$time0<=6, 1, 0)
lab.med$time1y<-ifelse(lab.med$time0>6 & lab.med$time0<=12, 1, 0)
lab.med$timeafter1y<-ifelse(lab.med$time0>12, 1, 0)

lab.med$timemonth<-round(lab.med$time0)
```

<br>

### 1.10 lab.med with baseline data
```{r}
compdat<-left_join(lab.med, base, by="PP187ID")
dim(compdat)

```

<br>


### 1.11 Check irregularities

```{r}
#compdat$valuen2[compdat$descriptio2=="Glucose nuchter"] %>% tapply(round(compdat$time0[compdat$descriptio2=="Glucose nuchter"]), mean, na.rm=T)
```
  + Seems like people stop using the medication in a long term?? the values regress back to the level before medication use.

<br>

##### 1.11.1 Remove observations who does not have any information before being prescribed for medication.
```{r}
removeid<-as.numeric()
for(i in unique(compdat$PP187ID)){
removeid[length(removeid)+1] <-ifelse(sum(compdat$time0 <=0 & compdat$PP187ID==i)==0,
                               i,
                              NA)
}
removeid<-removeid[!is.na(removeid)]
compdat<-compdat[-which(compdat$PP187ID %in% removeid), ]
```


<br>

##### 1.11.2 Rremoving t-1 before medication prescription to avoid random high.
```{r}
#Remove the last observations before the prescription
removerow<-as.numeric()
for(i in unique(compdat$PP187ID)){
  #return the row numbers that contain the last observations before the medication prescription
  
    removerow[length(removerow)+1] <- which(compdat$time0==sort(compdat$time0[compdat$time0 <=0 & compdat$PP187ID==i], decreasing = T)[1] & compdat$PP187ID==i)[1]
    removerow[length(removerow)+1] <- which(compdat$time0==sort(compdat$time0[compdat$time0 <=0 & compdat$PP187ID==i], decreasing = T)[1] & compdat$PP187ID==i)[2]
  }

removerow<-removerow[!is.na(removerow)]
compdat<-compdat[-removerow, ]
```

<br>


##### 1.11.3 Testrict data to 12m before & after prescription. To few observations (<8) outside the timeframe.
```{r}
#Save the complete data for later
compdat_ori<-compdat
compdat<-compdat_ori %>% filter(abs(time0)<=12)
compdat$timemonth<-round(compdat$time0)
```

<br>

##### 1.12 Final datasets
```{r}
compdat_hba1c<-compdat %>% filter(descriptio2=="HbA1c DCCT")
compdat_glu<-compdat %>% filter(descriptio2=="Glucose nuchter")


print("dimention of the complete set")
dim(compdat)

print("dimention of the hba1c data")
dim(compdat_hba1c)

print("dimention of the glucose data")
dim(compdat_glu)
#head(compdat)


```


<br>

<br>

# Step 2: Descriptive statistics

<br>

### 2.1 summarize the number of reapeated measurements per person/ mean values before and after presciption

```{r}
#To creat datasets to summarize

nrepeatsum<-function(data){

nrepeat<-data.frame()

for(i in data$PP187ID %>% unique){
  
  tmpdata<-data %>% filter(PP187ID==i)
  
  # ID
  nrepeat[i,1]<-i
  
  # total number of repeated measurements
  nrepeat[i,2]<-nrow(tmpdata)
  
  # number of repeated measurements after prescription 
  nrepeat[i,3]<-sum(tmpdata$time0>0)
  
  # Timing of the second last measurement before prescription
  nrepeat[i,4]<-sort(unique(tmpdata$time0[tmpdata$time0 <=0]), decreasing = T)[2]
  
  # Timing of the last measurement before prescription
  nrepeat[i,5]<-sort(unique(tmpdata$time0[tmpdata$time0 <=0]), decreasing = T)[1]
  
  # Value of the second last measurement before prescription
  nrepeat[i,6]<-tmpdata$valuen2[tmpdata$time0==sort(tmpdata$time0[tmpdata$time0 <=nrepeat[i,4]], decreasing = T)[1]][1]
  
  # Value of the last measurement before prescription
  nrepeat[i,7]<-tmpdata$valuen2[tmpdata$time0==sort(tmpdata$time0[tmpdata$time0 <=nrepeat[i,5]], decreasing = T)[1]][1]
  
  # Timing of the first measurement after prescription
  nrepeat[i,8]<-sort(unique(tmpdata$time0[tmpdata$time0 >0]), decreasing = F)[1]
  
  # Value of the first measurement after prescription
  nrepeat[i,9]<-tmpdata$valuen2[tmpdata$time0==sort(tmpdata$time0[tmpdata$time0 >nrepeat[i,8]], decreasing = F)[1]][1]
}
  colnames(nrepeat)<-c("PP187ID", 
                     "nrepeated", 
                     "n_after_med", 
                     "t_2nd_before",
                     "t_before",
                     "labval_2nd_before",
                     "labval_before",
                     "t_after",
                     "labval_after")

nrepeat<-subset(nrepeat, nrepeated != 0 & n_after_med != 0 )
nrepeat

}
```

<br>

  + **Summary data for HbA1c**
```{r}
nrepeat.hba1c<-nrepeatsum(compdat_hba1c)
#head(nrepeat.hba1c)
```

```{r}
print('number of repeated measurements for HbA1c (median/iqr)')
nrepeat.hba1c$nrepeated %>% median; 
nrepeat.hba1c$nrepeated %>% quantile(probs=c(0.25, 0.75))

print('Mean of the last measurement before the prescription (mean/sd)')
nrepeat.hba1c$labval_before %>% mean(na.rm=TRUE); 
nrepeat.hba1c$labval_before %>% sd(na.rm=TRUE)

print('Mean of the first measurement after the prescription (mean/sd)')
nrepeat.hba1c$labval_after %>% mean(na.rm=TRUE); 
nrepeat.hba1c$labval_after %>% sd(na.rm=TRUE)

```


<br>

  + **Summary data for glucose**
```{r}
nrepeat.glu<-nrepeatsum(compdat_glu) 
#head(nrepeat.glu)
```

```{r}
print('number of repeated measurements (median/iqr)')
nrepeat.glu$nrepeated %>% median; 
nrepeat.glu$nrepeated %>% quantile(probs=c(0.25, 0.75))

print('Mean of the last measurement before the prescription (mean/sd)')
nrepeat.glu$labval_before %>% mean(na.rm=TRUE); 
nrepeat.glu$labval_before %>% sd(na.rm=TRUE)

print('Mean of the first measurement after the prescription (mean/sd)')
nrepeat.glu$labval_after %>% mean(na.rm=TRUE); 
nrepeat.glu$labval_after %>% sd(na.rm=TRUE)

```


### 2.2 Make a baseline table
```{r}
#table1(~sexe+ 
#        leeftijd +
#        etnwhite +
#        eduh +
#        income +
#        bpsystgem +
#        bpdiastgem +
#        hypertensie+
#        lengtem +
#        gewichtm +
#        bmim +
#        vetpercentage +
#        middelomtrek +
#        heupomtrek +
#        Abdomen +
#        Vetpresent +
#        MSAT +
#        MVAT +
#        SSAT + 
#        SVAT +
#        VATSATratio +
#        ATPIIIglucose +
#        glucmet2bl+
#        nuchter +
#        devmeal +
#        glucose1 +
#        glucose2 +
#        glucose3 +
#        Insuline_r1 +
#        insuline_r2 +
#        insuline_r3 +
#        homa1B +
#        homa1IR +
#        HBA1C +
#        choltot1 +
#        trig1 +
#        trig2 +
#        trig3 + 
#        hdlc1 +
#        fldl1 +
#        ALBM1 +
#        screat +
#        VitD25OH1_cal +
#        calcialb +
#        calcium1 +
#        natrium +
#        kalium +
#        urizuur +
#        medC10liplow +
#        medA10gluclow +
#        medHypertension +
#        medN06A +
#        medN05A +
#        medN05B +
#        medH03A +
#        medL02A +
#        medL02B,
#       data=base.comp)

table1(~sexe+ 
        leeftijd +
        eduh +
        hypertensie+
        bmim +
        glucose1 +
        Insuline_r1 +
        homa1IR +
        HBA1C +
        choltot1 +
        trig1 +
        hdlc1 +
        fldl1 +
        medC10liplow +
        medA10gluclow +
        medHypertension,
       data=base.comp)

```

<br>

### 2.3 Box plots: values at eath time point (in month)
  + To Olaf: are there any values unrealistic?
```{r}

give.n <- function(x){
  return(c(y = 5, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}


box_hba1c <- compdat_hba1c %>% ggplot(aes(factor(timemonth), valuen2)) + 
              geom_boxplot() +
              ylim(5,13) +
              stat_summary(fun.data = give.n, geom = "text", fun = median, position = position_dodge(width = 0.75)) +
  ylab("HbA1c DCCT") +
  xlab("Time in month")


box_glu <- compdat_glu %>% ggplot(aes(factor(timemonth), valuen2)) + 
              geom_boxplot() +
              ylim(5,13) +
              stat_summary(fun.data = give.n, geom = "text", fun = median,
                                position = position_dodge(width = 0.75)) +
  ylab("Glucose") +
  xlab("Time in month")


box_hba1c  ; box_glu
```

<br>

### 2.3 Spaghetti plots for every participants
  + **HbA1c DCCT**
```{r}
plotlist_hba1c<-list()

for(i in 1:15){
#spaghetti plots per 15 individuals
  plot<-compdat_hba1c[compdat_hba1c$PP187ID %in% unique(compdat_hba1c$PP187ID)[((i-1)*15+1):(15*i)],] %>% 
        ggplot(aes(x = time0, y = valuen2, group= PP187ID)) + 
        geom_line() +
        xlim(-12, 12) +
        aes(colour = factor(PP187ID)) +
        labs(x="Number of months (Time0: prescription date)", y="HbA1c DCCT [%]")
plotlist_hba1c[[i]]<-plot
}

plotlist_hba1c
```

<br>

  + **Glucose nucther**
```{r}
plotlist_glu<-list()

for(i in 1:15){
#spaghetti plots per 15 individuals
  plot<- compdat_glu[compdat_glu$PP187ID %in% unique(compdat_glu$PP187ID)[((i-1)*15):(15*i)],] %>% 
         ggplot(aes(x = time0, y = valuen2, group= PP187ID)) + 
         geom_line() +
         aes(colour = factor(PP187ID)) +
         labs(x="Number of months (Time0: prescription date)", y="Glucose [mmol/L]")
plotlist_glu[[i]]<-plot
}

plotlist_glu
```


<br>

<br>

# Step 3: fitting LMMs

<br>

### 3.1 HbA1c

<br>

#### Model 0 HbA1c: intercept only
```{r}
hba1c.m0<- lmer(valuen2~ (1 |PP187ID),
                data=compdat_hba1c)

summary(hba1c.m0)
```

<br>

#### Model 1 HbA1c: medication use as a fixed and a random effect
```{r}
hba1c.m1<- lmer(valuen2~ (1+medindi|PP187ID) + medindi,
                data=compdat_hba1c)

summary(hba1c.m1)

```

<br>

#### Model 2 HbA1c: additional covariates
```{r}
hba1c.m2<- lmer(valuen2~ (1+medindi|PP187ID) + medindi + sexe + leeftijd + bmim,
                data=compdat_hba1c)

summary(hba1c.m2)
```

<br>

#### Model 3 HbA1c: interaction between time * medication use
```{r}
hba1c.m3<- lmer(valuen2~ (1+medindi|PP187ID) + medindi*time0 + sexe + leeftijd + bmim,
                data=compdat_hba1c)

summary(hba1c.m3)
```

<br>

```{r}
#Anova for myself
#anova(lmm.hba1c.m0, lmm.hba1c.m1, lmm.hba1c.m2, lmm.hba1c.m3)
```

<br>

#### Categorial time model
```{r}
##### Use these #####
#lab.med$timebefore_6m<-ifelse(lab.med$time0 <= -6, 1, 0)
#lab.med$time_6m<-ifelse(lab.med$time0 > -6 & lab.med$time0 <= -3, 1, 0)
#lab.med$time_3m<-ifelse(lab.med$time0 > -3 & lab.med$time0 <= 0, 1, 0)
#lab.med$time3m<-ifelse(lab.med$time0>0 & lab.med$time0<=3, 1, 0)
#lab.med$time6m<-ifelse(lab.med$time0>3 & lab.med$time0<=6, 1, 0)
#lab.med$time1y<-ifelse(lab.med$time0>6 & lab.med$time0<=12, 1, 0)


hba1c.m4<- lmer(valuen2~ (1+medindi|PP187ID) +
                time_6m + time_3m + time3m + time6m + time1y + sexe + leeftijd + bmim,
                data=compdat_hba1c)

summary(hba1c.m4)
```


<br>

<br>

### 3.2 Glucose

<br>

#### Model 0 glucose: intercept only
```{r}
glu.m0<- lmer(valuen2~ (1|PP187ID),
              data=compdat_glu)

summary(glu.m0)
```

<br>

#### Model 1 glucose: medication use as a fixed and a random effect
```{r}
glu.m1<- lmer(valuen2~ (1+medindi|PP187ID) + medindi,
              data=compdat_glu)

summary(glu.m1)
```

<br>

#### Model 2 HbA1c: additional covariates
```{r}
glu.m2<- lmer(valuen2~ (1+medindi|PP187ID) + medindi + sexe + leeftijd + bmim,
              data=compdat_glu)

summary(glu.m2)
```

<br>

#### Model 3 HbA1c: interaction between time * medication use
```{r}
glu.m3<- lmer(valuen2~ (1+medindi|PP187ID) + medindi*time0 + sexe + leeftijd + bmim,
              data=compdat_glu)

summary(glu.m3)
```


<br>

#### Categorical time model
```{r}
##### Use these #####
#lab.med$timebefore_6m<-ifelse(lab.med$time0 <= -6, 1, 0)
#lab.med$time_6m<-ifelse(lab.med$time0 > -6 & lab.med$time0 <= -3, 1, 0)
#lab.med$time_3m<-ifelse(lab.med$time0 > -3 & lab.med$time0 <= 0, 1, 0)
#lab.med$time3m<-ifelse(lab.med$time0>0 & lab.med$time0<=3, 1, 0)
#lab.med$time6m<-ifelse(lab.med$time0>3 & lab.med$time0<=6, 1, 0)
#lab.med$time1y<-ifelse(lab.med$time0>6 & lab.med$time0<=12, 1, 0)
glu.m4<- lmer(valuen2~ (1+medindi|PP187ID) + 
              time_6m + time_3m + time3m + time6m + time1y + sexe + leeftijd + bmim,
              data=compdat_glu)

summary(glu.m4)

```


<br>

### 3.3 Sumamry table for categorical models 
```{r}

summarytab<-function(results){
tabs<-list()  

#Random effects
raneff.name<-c("Intercept",
          "Medication use",
          "Intercept & Medication use",
          "Residual")

resvar<-results %>% VarCorr %>% as.data.frame 
raneff.covar<-resvar[,4] %>% round(2)
raneff.cosd<- resvar[,5] %>% round(2)
raneff.NA<-rep(NA, length(raneff.name))
raneff.dat<-data.frame(raneff.name, raneff.covar, raneff.cosd, raneff.NA)
colnames(raneff.dat)<-c("Random effects",
                        "Covariance",
                        "SD",
                        " ")

fixeff.name<-c("Intercept",
          "3m - 6m before prescription",
          "0 - 3m before prescription",
          "0 - 3m after prescription",
          "3m - 6m after prescription",
          "6m - 1y after prescription",
          "Gender (0=Male, 1=female)",
          "Age at the first visit",
          "BMI at the first visit")

resest<-results %>% summary %>% coef %>% round(2)
fixeff.est<- resest[,1]
fixeff.se<- resest[,2]
fixeff.tval<-resest[,3]
fixeff.dat<-data.frame(fixeff.name, fixeff.est, fixeff.se, fixeff.tval, row.names = )
colnames(fixeff.dat)<-c("Fixed effects",
                        "Estimate",
                        "SE",
                        "t-val")


tabs[[1]]<-gt(data=raneff.dat)
tabs[[2]]<-gt(data=fixeff.dat)
tabs
}
```

<br>

#### HbA1c summary
```{r}
summarytab(hba1c.m4)
```

<br>

#### Glucose summary
```{r}
summarytab(glu.m4)
```

<br>

<br>

# Step 4: fitting with splines 

<br>

### 4.1 HbA1c
```{r}
hba1c.sp<- lmer(valuen2~ (1+medindi|PP187ID) + rcs(time0,3) + medindi + medindi:rcs(time0,3),
              data=compdat_hba1c)


#hba1c.sp
```


```{r}
predicted<-hba1c.sp %>% predict(compdat_hba1c, re.form=NA)


plot(compdat_hba1c$time0, predicted, 
     xlab = "Number of months (Time0: prescription date)",
     ylab = "Predicted HbA1c [%]")

```



### 4.2 HbA1c

```{r}
glu.sp<- lmer(valuen2~ (1+medindi|PP187ID) + rcs(time0,3) + medindi + medindi:rcs(time0,3),
              data=compdat_glu)

predicted<-glu.sp %>% predict(compdat_glu, re.form=NA)
```


```{r}
plot(compdat_glu$time0, predicted, 
     xlab = "Number of months (Time0: prescription date)",
     ylab = "Predicted glucose [mmol/L]")

```


<br>

<br>



-------------------------------------------------------------------------------------------------------------------------------
***CONTINUE***

# Step 5: medication information <DATA REMOVED>

---------------------------------------------------------------------------------------------------------------------------
***Something to think of...***

1) Some people have multiple prescription drugs targeting glucose-related measurements, stopped using drugs, change their prescription. 

2) Ask Olaf: we noticed that last measurements before the prescription are high. Could be random high? -> solved

3) I have not added baseline measurement to the repeated measurement yet. Still waiting for the first NEO visits dates.

3) Double check the glucose measurements can indeed be combined into one.

4) Sensitivity analyses??


